---
title: "Lab3"
author: "Tessa Mendoza"
date: "9/22/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
dat <- readxl::read_xls('GSS.xls')
library(tidyverse)
```

1)
```{r warning=FALSE}
dat <- slice(dat, 1:(n()-2))
dat <- dat %>% rename('education' = `Highest year of school completed`,
                      'age' = `Age of respondent`, party = `Political party affiliation`, 'marriage' = `Marital status`, 'happy' = `General happiness`)

dat <- dat %>%
  mutate(education = as.numeric(education),
         age = as.numeric(age)) %>%
  mutate(party = factor(party))

meanEdu <- dat %>%
  filter(!is.na(education)) %>%
  group_by(party) %>%
  summarize(mean_education = mean(education)) %>%
  arrange(desc(mean_education))

meanEdu
```
While the mean number of years is similar between each political party group and ranges from 10-13 years, there is one noticable pattern: those who do not know their political party tend to have a lower amount of education (on average, through part of high school). This suggests that those with higher education tend to have a more distinguished indentification with a political party. 


2)
```{r}
betterEdu <- dat %>%
  filter(!is.na(education)) %>%
  group_by(party) %>%
  filter(education > median(education, na.rm = TRUE)) %>%
  ungroup()
str(betterEdu)
```

3)
```{r}
happy <- dat %>%
  mutate(marriage = factor(marriage)) %>%
  filter(happy == "Very happy" | happy == "Pretty happy")

happytable <- happy %>%
  group_by(marriage) %>%
  summarize(numberHappy = n()) %>%
  arrange(desc(numberHappy))
happytable

ggplot(happy, aes(x = happy, fill = marriage)) + geom_bar(position = 'stack')

```
4)
```{r}
partyaff <- dat %>%
  group_by(party, marriage) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

partyaff

partyaffhappy <- dat %>%
  filter(happy == "Very happy" | happy == "Pretty happy")

ggplot(partyaffhappy, aes(x = party, fill = marriage)) + coord_flip() + geom_bar(position = 'stack') + ggtitle("Party and Marriage Distribution Among Happy People")

```


5)
```{r}
dat <- dat %>%
  mutate(education = as.numeric(education),
         age = as.numeric(age))

dat <- dat %>%
  mutate(party = factor(party)) %>%
  mutate(marriage = factor(marriage))

 
marry <- dat %>%
  filter(!is.na(education)) %>%
  group_by(marriage) %>%
  summarize(mean_education = mean(education)) %>%
  arrange(desc(mean_education))
marry
dat <- dat %>%
  filter(!is.na(education))
ggplot(dat, aes(x = marriage, y = education)) + geom_boxplot() + stat_summary(fun.y = mean, geom = 'point', color = "red", size = 4)

```

There is a general relationship between marriage status and mean education. Never married, divorced, and married persons had a higher mean education status compared to separated and widowed persons. This could suggest that being widowed or separated are events that keep a person from continuing his or her education. Perhaps, separated and widowed people are more likely to be single parents, and therefore have to overcome higher obstacles to attend school.

6) 
```{r}
ageHappy <- dat %>%
  mutate(happy = factor(happy)) %>%
  filter(!is.na(age)) %>%
  group_by(happy) %>%
  summarize(mean_age = mean(age))
ageHappy

ageHappyPlot<- dat %>%
  filter(!is.na(age))
ggplot(ageHappyPlot, aes(x = happy, y = age)) + geom_boxplot()

```

As a whole, I found this plot to be very interesting. From this simple analysis, it seems as if there is not a strong correlation between age and happiness. In fact, the mean value is very similar across all happiness ratings. However, there is a slight increase in mean age for those who don't know how to rank their happiness. As shown in the boxplot, the range of values occurs at a higher age than the other happiness values. This suggests that younger people (under 30) are more strongly decisive in whether or not they are happy.
