---
title: "Lab3"
author: "Tessa Mendoza"
date: "9/22/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
dat <- readxl::read_xls('GSS.xls')
library(tidyverse)
```

1)
```{r warning=FALSE}
dat <- slice(dat, 1:(n()-2))
dat <- dat %>% rename('education' = `Highest year of school completed`,
                      'age' = `Age of respondent`, party = `Political party affiliation`, 'marriage' = `Marital status`)

dat <- dat %>%
  mutate(education = as.numeric(education),
         age = as.numeric(age)) %>%
  mutate(party = factor(party))

meanEdu <- dat %>%
  filter(!is.na(education)) %>%
  group_by(party) %>%
  summarize(mean_education = mean(education)) %>%
  arrange(desc(mean_education))

meanEdu
```
While the mean number of years is similar between each political party group and ranges from 10-13 years, there is one noticable pattern: those who do not know their political party tend to have a lower amount of education (on average, through part of high school). This suggests that those with higher education tend to have a more distinguished indentification with a political party. 


2)
```{r}
betterEdu <- dat %>%
  filter(!is.na(education)) %>%
  group_by(party) %>%
  filter(education > median(education, na.rm = TRUE)) %>%
  ungroup()
str(betterEdu)
```

3)
```{r}
happy <- dat %>%
  mutate(marriage = factor(marriage)) %>%
  rename(happy = `General happiness`) %>%
  #mutate(happy = factor(happy)) %>%
  filter(happy == "Very happy" | happy == "Pretty happy")

happytable <- happy %>%
  group_by(marriage) %>%
  summarize(numberHappy = n()) %>%
  arrange(desc(numberHappy))
happytable
ggplot(happy, aes(x = marriage)) + geom_bar()
```


5)
```{r}
dat <- dat %>%
  mutate(education = as.numeric(education),
         age = as.numeric(age))

dat <- dat %>%
  mutate(party = factor(party)) %>%
  mutate(marriage = factor(marriage))

 
marry <- dat %>%
  filter(!is.na(education)) %>%
  group_by(marriage) %>%
  summarize(mean_education = mean(education)) %>%
  arrange(desc(mean_education))
marry
ggplot(marry, aes(x = marriage, y = mean_education)) + geom_bar(stat = 'identity')
```

There is a general relationship between marriage status and mean education. Never married, divorced, and married persons had a higher mean education status compared to separated and widowed persons. This could suggest that being widowed or separated are events that keep a person from continuing his or her education. Perhaps, separated and widowed people are more likely to be single parents, and therefore have to overcome higher obstacles to attend school.